---
title: "Developing Data Products - Week 2 Peer-graded Assignment: R Markdown and Leaflet"
author: "Lisa Halmschlager"
date: "6/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Create a web page using R Markdown that features a map created with Leaflet.

Host your webpage on either GitHub Pages, RPubs, or NeoCities.

Your webpage must contain the date that you created the document, and it must contain a map created with Leaflet. We would love to see you show off your creativity!

## About Leaflet
Leaflet is a JavaScript library, and then there is an associated Leaflet package for R that makes use of this JavaScript library to create interactive maps within the R environment. 

The `leaflet()` function creates a map widget that can be stored in a variable so that the map can be modify later on. One can
add features to the map using the pipe operator (`%>%`) just like in dplyr. 

The `addTiles()` function adds mapping data from [Open Street Map](http://www.openstreetmap.org/).

It is possible to add markers to the map using the `addMarkers()` function by specifying the longitude and latitude.
One can make own markers from `.png` files or add popups for each marker (string of plain text, or HTML to be rendered inside of each popup).

Leaflet allows you to plot clusters of markers using `addMarkers(clusterOptions = markerClusterOptions())`. When zooming in to each cluster, the clusters will separate until the individual markers can be seen.


```{r leaflet, echo=FALSE, warning=FALSE}
# install.packages("leaflet")
library(leaflet)
```

## Assignment: Map of CEU Campus

Since I am going to study Business Analytics at Central European University from September 2019, I decided to map out the different buildings in Budapest that belong to the university:

### CEU Buildings:
1. Nador utca 15 is the key entrance to the University and is home to the CEU Auditorium, the CEU Library, classrooms, a cafe and a restaurant. The building opened in fall 2016 with neighboring Nador utca 13. The four interconnected buildings house classrooms, academic departments, administrative units, and student study spaces.
2. Zrinyi utca 14 houses some classrooms, academic and administrative units.
3. Oktober 6. utca 7. This building houses classrooms, the Oktober Hall event space and several academic units.
4. Vigyazo Ferenc utca 2 houses academic units and faculty offices.
5. Jozsef Attila utca 24. This building houses administrative units.
6. CEU Press
7. Vera and Donald Blinken Open Society Archives

```{r map2, message=FALSE}
# data frame with coordinates of all CEU buildings 
df <- data.frame(lat = c(47.5012048, 47.5, 47.502547, 47.501003, 47.499130, 47.500644, 47.502647),
                 lng = c(19.0499434, 19.05, 19.050616, 19.047830, 19.053587, 19.050905, 19.053404),
                 col = c("blue", "blue","blue", "yellow","yellow","yellow","yellow"), 
                 stringsAsFactors = FALSE)

# study logo
StudyIcon <- makeIcon(
  iconUrl = "https://cdn.onlinewebfonts.com/svg/img_534893.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

# popups
CEUSites <- c(
  "<a href='https://www.ceu.edu//'>Nador utca 15</a>",
  "<a href='https://www.ceu.edu/'>Zrinyi utca 14s</a>",
  "<a href='https://www.ceu.edu/'>Oktober 6. utca 7</a>",
  "<a href='https://www.ceu.edu/'>Vigyazo Ferenc utca 2</a>",
  "<a href='https://www.ceu.edu/'>Jozsef Attila utca 24</a>",
  "<a href='http://ceupress.com/'>CEU Press</a>",
  "<a href='http://www.osaarchivum.org/'>Vera and Donald Blinken Open Society Archives</a>"
)

df %>%  
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(icon = StudyIcon, popup = CEUSites)
```


```{r no_piping, echo=FALSE, eval=FALSE}
# above: piping notation
# identical code:
my_map = leaflet()
my_map = addTiles(my_map)
my_map = addMarkers(my_map, lat=39.2980803, lng=-76.5898801, popup="Jeff Leek's Office")


df %>%  
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(icon = StudyIcon, popup = CEUSites) %>%
  addLegend(labels = c("Classrooms", "Other Sites"), colors = c("blue", "yellow"))
```
